---
title: "Enrichment analysis"
author: "Jill Ashey"
date: "`r Sys.Date()`"
output: html_document
---

This script will perform functional enrichment analysis with GO and KEGG terms using the [identified DEGs](https://github.com/fscucchia/Hawaii2022_pH_Temp_Mcap_Pacu/blob/main/scripts/DESeq2.Rmd) from the Pacuta 2022 HI experiment. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load libraries
library("DESeq2")
library("tidyverse")
library("dplyr")
library("pheatmap")
library("RColorBrewer")
library("genefilter")
library("ggplot2")
library("gplots")
library("limma")
library("spdep") 
library("adegenet") 
library("goseq")
library("gridExtra")
library("clusterProfiler")
library("DataCombine")
library("VennDiagram")
```

Read in all expressed genes (poverA = 0.75, 5)
```{r}
filt_counts <- read.csv("output/DEG/filtered_counts.csv")
colnames(filt_counts)[1] <- "gene_id"
```

Read in DEGs
```{r}
degs <- read.csv("output/DEG/Unique_DEGs.csv")
colnames(degs)[1] <- "gene_id"
```

Calculate read lengths from Pacuta V2 [cds file](http://cyanophora.rutgers.edu/Pocillopora_acuta/). I did this on my local computer using the code below. 
```{bash}
cd /Users/jillashey/Desktop/GFFs/pacuta/V2

awk 'BEGIN{FS="[> ]"} /^>/{val=$2;next}  {print val,length($0)}' Pocillopora_acuta_HIv2.genes.cds.fna > gene_length.txt
```

Read in gene lengths 
```{r}
length <- read.csv("output/DEG/gene_length.txt", header = F)

length <- length %>%
  separate(V1, into = c("gene_id", "length"), sep = " ", remove = FALSE) %>%
  select(-V1)
```

Merge length data with filtered count data
```{r}
length_merge <- merge(length, filt_counts, by = "gene_id")
```

GOseq requires a vector of all genes, all differentially expressed genes, and gene lengths. Make vectors 
```{r}
# Make DEG vector
DEG <- length_merge[length_merge$gene_id %in% degs$gene_id, ]
DEG_names <- as.vector(DEG$gene_id)
gene_vector <- as.integer(length_merge$gene_id%in%DEG_names)
names(gene_vector) <- length_merge$gene_id

# Make ID vector
ID_vector <- length_merge$gene_id

# Make length vector
length_vector <- as.numeric(length_merge$length)
```

Calculate probability weighting function 
```{r}
DEG.pwf <- nullp(gene_vector, ID_vector, bias.data = length_vector)
```

Code above giving me this warning: Warning: initial point very close to some inequality constraintsWarning: collapsing to unique 'x' values

Read in GO annotation information
```{r}
annot <- read.delim("Metadata/Pocillopora_acuta_HIv2.genes.EggNog_results.txt", header = T)
```

This annot file has GO, Kegg, eggNog and Pfam annotation information. I'm going to start analyzing the GO terms first.

Select GO columns and split so that there is only one GO term per row
```{r}
annot_GO <- annot %>%
  select(X.query, GOs) %>%
  mutate_all(~if_else(. == "-", NA_character_, .)) %>%
  separate_rows(GOs, sep = ",") %>%
  rename(gene_id = X.query,
         GO.ID = GOs)

# Remove leading and trailing whitespaces
annot_GO$GO.ID <- trimws(annot_GO$GO.ID)

# Replace NAs with unknown
annot_GO$GO.ID <- replace_na(annot_GO$GO.ID, "unknown")

# Set gene and go ids as factors 
annot_GO$gene_id <- as.factor(annot_GO$gene_id)
annot_GO$GO.ID <- as.factor(annot_GO$GO.ID)

# Select only unique rows
annot_GO <- unique(annot_GO)

# Look at unique gene and go ids 
length(unique(annot_GO$gene_id))
length(unique(annot_GO$GO.ID))

# Remove rows with unknowns
annot_GO <- annot_GO %>%
  filter(GO.ID != "unknown")
```

Perform GOSeq
```{r}
GO.wall<-goseq(DEG.pwf, ID_vector, gene2cat=annot_GO, method="Wallenius", use_genes_without_cat=T)
```

Getting this error - Error: C stack usage  7970448 is too close to the limit. I think R doesn't have enough memory?? Need to troubleshoot this further. 

https://stackoverflow.com/questions/14719349/error-c-stack-usage-is-too-close-to-the-limit 







