---
title: "Enrichment analysis"
author: "Jill Ashey"
date: "`r Sys.Date()`"
output: html_document
---

This script will perform functional enrichment analysis with GO and KEGG terms using the [identified DEGs](https://github.com/fscucchia/Hawaii2022_pH_Temp_Mcap_Pacu/blob/main/scripts/DESeq2.Rmd) from the Pacuta 2022 HI experiment. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load libraries
library("DESeq2")
library("tidyverse")
library("dplyr")
library("pheatmap")
library("RColorBrewer")
library("genefilter")
library("ggplot2")
library("gplots")
library("limma")
library("spdep") 
library("adegenet") 
library("goseq")
library("gridExtra")
library("clusterProfiler")
library("DataCombine")
library("VennDiagram")
library(rtracklayer)
#library(rrvgo)
library(GO.db)
library(tidyr)
library(forcats)

#devtools::install_github("krlmlr/ulimit")
#library(ulimit)

#ulimit::memory_limit(7970592)
```

Read in all expressed genes (poverA = 0.75, 5)
```{r}
filt_counts <- read.csv("../output/DEG/filtered_counts.csv")
colnames(filt_counts)[1] <- "gene_id"

str(filt_counts)
```

Read in DEGs
```{r}
degs <- read.csv("../output/DEG/Unique_DEGs.csv")
colnames(degs)[1] <- "gene_id"

str(degs)
```

Calculate read lengths from Pacuta V2 [cds file](http://cyanophora.rutgers.edu/Pocillopora_acuta/). I did this on my local computer using the code below. 
```{bash}
cd /Users/jillashey/Desktop/GFFs/pacuta/V2

awk 'BEGIN{FS="[> ]"} /^>/{val=$2;next}  {print val,length($0)}' Pocillopora_acuta_HIv2.genes.cds.fna > gene_length.txt
```

Read in gene lengths 
```{r}
length <- read.csv("../output/DEG/gene_length.txt", header = F)

length <- length %>%
  separate(V1, into = c("gene_id", "length"), sep = " ", remove = FALSE) %>%
  dplyr::select(-V1)

str(length)
```

Merge length data with filtered count data
```{r}
length_merge <- merge(length, filt_counts, by = "gene_id")

str(length_merge)
```

GOseq requires a vector of all genes, all differentially expressed genes, and gene lengths. Make vectors 
```{r}
# Make DEG vector
DEG <- length_merge[length_merge$gene_id %in% degs$gene_id, ]
DEG_names <- as.vector(DEG$gene_id)
gene_vector <- as.integer(length_merge$gene_id%in%DEG_names)
names(gene_vector) <- length_merge$gene_id

# Make ID vector
ID_vector <- as.vector(length_merge$gene_id)

# Make length vector
length_vector <- as.numeric(length_merge$length)
```

Calculate probability weighting function 
```{r}
DEG.pwf <- nullp(gene_vector, ID_vector, bias.data = length_vector)

str(DEG.pwf)
```

Code above giving me this warning: Warning: initial point very close to some inequality constraintsWarning: collapsing to unique 'x' values

Read in GO annotation information
```{r}
annot <- read.delim("../Metadata/Pocillopora_acuta_HIv2.genes.EggNog_results.txt", header = T)
```

This annot file has GO, Kegg, eggNog and Pfam annotation information. I'm going to start analyzing the GO terms first.

Select GO columns and split so that there is only one GO term per row
```{r}
annot_GO <- annot %>%
  dplyr::select(X.query, GOs) %>%
  mutate_all(~if_else(. == "-", NA_character_, .)) %>%
  separate_rows(GOs, sep = ",") %>%
  rename(gene_id = X.query,
         GO.ID = GOs)

# Remove leading and trailing whitespaces
annot_GO$GO.ID <- trimws(annot_GO$GO.ID)

# Replace NAs with unknown
annot_GO$GO.ID <- replace_na(annot_GO$GO.ID, "unknown")

# Set gene and go ids as factors 
annot_GO$gene_id <- as.factor(annot_GO$gene_id)
annot_GO$GO.ID <- as.factor(annot_GO$GO.ID)

# Select only unique rows
annot_GO <- unique(annot_GO)

# Look at unique gene and go ids 
length(unique(annot_GO$gene_id))
length(unique(annot_GO$GO.ID))

# Remove rows with unknowns
# annot_GO <- annot_GO %>%
#   filter(GO.ID != "unknown")
```

Remove the annotation information for genes not present in our data 
```{r}
filt_annot_GO <- subset(annot_GO, gene_id %in% filt_counts$gene_id)

str(filt_annot_GO)

# Convert filt_annot_GO from tibble to df 
filt_annot_GO <-as.data.frame(filt_annot_GO)
```


Perform GOSeq
```{r}
GO.wall<-goseq::goseq(DEG.pwf, ID_vector, gene2cat=filt_annot_GO, method="Wallenius", use_genes_without_cat=T)

GO <- GO.wall[order(GO.wall$over_represented_pvalue),]
write.csv(GO, file = "../output/functional_enrichment/GOSeq_GO_all.csv")

class(GO.wall)
head(GO.wall)
tail(GO.wall)
nrow(GO.wall)
```

Worked! 

Filter by significantly enriched GO terms (p < 0.05)
```{r}
GO_05 <- GO %>%
  dplyr::filter(over_represented_pvalue<0.05) %>%
        dplyr::arrange(., ontology, over_represented_pvalue)

write.csv(GO_05, file = "../output/functional_enrichment/GOSeq_GO_05.csv")
```

Plot all ontologies (BP, CC, MF) and order by p-value 
```{r}
GO_05_plot <- GO_05 %>% drop_na(ontology) %>% mutate(term = fct_reorder(term, numDEInCat)) %>%
  mutate(term = fct_reorder(term, ontology)) %>%
  ggplot( aes(x=term, y=numDEInCat) ) +
  geom_segment( aes(x=term ,xend=term, y=0, yend=numDEInCat), color="grey") +
  geom_text(aes(label = over_represented_pvalue), hjust = -1, vjust = 0, size = 2) +
  geom_point(size=1, aes(colour = ontology)) +
  coord_flip() +
  ylim(0,305) +
  theme(
    panel.grid.minor.y = element_blank(),
    panel.grid.major.y = element_blank(),
    legend.position="bottom"
  ) +
  xlab("") +
  ylab("") +
  theme_bw() + #Set background color 
  theme(panel.border = element_blank(), # Set border
        panel.grid.major = element_blank(), #Set major gridlines
        panel.grid.minor = element_blank(), #Set minor gridlines
        axis.line = element_line(colour = "black"), #Set axes color
        plot.background=element_blank()); GO_05_plot #Set the plot background #set title attributes

ggsave("../output/functional_enrichment/GOSeq_GO_05.pdf", GO_05_plot, width = 20, height = 10)
ggsave("../output/functional_enrichment/GOSeq_GO_05.png", GO_05_plot, width = 20, height = 10)
```

Plot ontologies (BP, CC, MF) separately and order by p-value 
```{r}
## BP 
GO_05_plot_BP <- GO_05 %>% 
  drop_na(ontology) %>% 
  filter(ontology == "BP") %>%
  mutate(term = fct_reorder(term, numDEInCat)) %>%
  #mutate(term = fct_reorder(term)) %>%
  ggplot( aes(x=term, y=numDEInCat) ) +
  geom_segment( aes(x=term ,xend=term, y=0, yend=numDEInCat), color="grey") +
  geom_text(aes(label = over_represented_pvalue), hjust = -1, vjust = 0, size = 2) +
  geom_point(size=1) +
  coord_flip() +
  ylim(0,305) +
  theme(
    panel.grid.minor.y = element_blank(),
    panel.grid.major.y = element_blank(),
    legend.position="bottom"
  ) +
  xlab("") +
  ylab("") +
  theme_bw() + #Set background color 
  theme(panel.border = element_blank(), # Set border
        panel.grid.major = element_blank(), #Set major gridlines
        panel.grid.minor = element_blank(), #Set minor gridlines
        axis.line = element_line(colour = "black"), #Set axes color
        plot.background=element_blank()); GO_05_plot_BP #Set the plot background #set title attributes

ggsave("../output/functional_enrichment/GOSeq_GO_05_BP.pdf", GO_05_plot_BP, width = 30, height = 10)
ggsave("../output/functional_enrichment/GOSeq_GO_05_BP.png", GO_05_plot_BP, width = 30, height = 10)

## CC
GO_05_plot_CC <- GO_05 %>% 
  drop_na(ontology) %>% 
  filter(ontology == "CC") %>%
  mutate(term = fct_reorder(term, numDEInCat)) %>%
  #mutate(term = fct_reorder(term)) %>%
  ggplot( aes(x=term, y=numDEInCat) ) +
  geom_segment( aes(x=term ,xend=term, y=0, yend=numDEInCat), color="grey") +
  geom_text(aes(label = over_represented_pvalue), hjust = -1, vjust = 0, size = 2) +
  geom_point(size=1) +
  coord_flip() +
  ylim(0,305) +
  theme(
    panel.grid.minor.y = element_blank(),
    panel.grid.major.y = element_blank(),
    legend.position="bottom"
  ) +
  xlab("") +
  ylab("") +
  theme_bw() + #Set background color 
  theme(panel.border = element_blank(), # Set border
        panel.grid.major = element_blank(), #Set major gridlines
        panel.grid.minor = element_blank(), #Set minor gridlines
        axis.line = element_line(colour = "black"), #Set axes color
        plot.background=element_blank()); GO_05_plot_CC #Set the plot background #set title attributes

ggsave("../output/functional_enrichment/GOSeq_GO_05_CC.pdf", GO_05_plot_CC, width = 30, height = 10)
ggsave("../output/functional_enrichment/GOSeq_GO_05_CC.png", GO_05_plot_CC, width = 30, height = 10)

## MF
GO_05_plot_MF <- GO_05 %>% 
  drop_na(ontology) %>% 
  filter(ontology == "MF") %>%
  mutate(term = fct_reorder(term, numDEInCat)) %>%
  #mutate(term = fct_reorder(term)) %>%
  ggplot( aes(x=term, y=numDEInCat) ) +
  geom_segment( aes(x=term ,xend=term, y=0, yend=numDEInCat), color="grey") +
  geom_text(aes(label = over_represented_pvalue), hjust = -1, vjust = 0, size = 2) +
  geom_point(size=1) +
  coord_flip() +
  ylim(0,305) +
  theme(
    panel.grid.minor.y = element_blank(),
    panel.grid.major.y = element_blank(),
    legend.position="bottom"
  ) +
  xlab("") +
  ylab("") +
  theme_bw() + #Set background color 
  theme(panel.border = element_blank(), # Set border
        panel.grid.major = element_blank(), #Set major gridlines
        panel.grid.minor = element_blank(), #Set minor gridlines
        axis.line = element_line(colour = "black"), #Set axes color
        plot.background=element_blank()); GO_05_plot_MF #Set the plot background #set title attributes

ggsave("../output/functional_enrichment/GOSeq_GO_05_MF.pdf", GO_05_plot_MF, width = 30, height = 10)
ggsave("../output/functional_enrichment/GOSeq_GO_05_MF.png", GO_05_plot_MF, width = 30, height = 10)
```

Merge significant GO terms and annot list together 
```{r}
GO_genes <- annot_GO %>%
  inner_join(GO_05, by = c("GO.ID" = "category"))

length(unique(GO_genes$gene_id)) # 3786 genes with over-represented GO terms 
```

Read in DEGs and merge that data with significant GO terms 
```{r}
deg <- read.csv("../output/DEG/Unique_DEGs.csv") %>%
  rename("gene_id" = "X")

# Merge DEGs and biomin gene information
GO_deg <- GO_genes %>%
  inner_join(deg, by = "gene_id")

length(unique(GO_deg$gene_id)) # 178 DEGs with over-represented GO terms 

write.csv(GO_deg, file = "../output/functional_enrichment/DEG_GO.csv")
```

How does this relate to our DEG comparisons? Merge high v control DEGs with GO information
```{r}
## High vs control treatment comparison 
h_v_c <- read.csv("../output/DEG/High_v_Control_DEGs.csv") %>%
  rename("gene_id" = "X")

# Merge DEGs and GO information and remove extra columns
GO_h_v_c <- h_v_c %>%
  inner_join(GO_deg, by = "gene_id") %>%
  dplyr::select(-c("A.y", "B.y", "C.y", "D.y", "E.y", "F.y", "G.y", "H.y", "L.y", "M.y", "N.y"))

length(unique(GO_h_v_c$gene_id))

write.csv(GO_h_v_c, file = "../output/functional_enrichment/High_v_Control_GO.csv")
```

Merge Mid v control DEGs with GO information
```{r}
## Mid vs control treatment comparison 
m_v_c <- read.csv("../output/DEG/Mid_v_Control_DEGs.csv") %>%
  rename("gene_id" = "X")

# Merge DEGs and GO information and remove extra columns
GO_m_v_c <- m_v_c %>%
  inner_join(GO_deg, by = "gene_id") %>%
  dplyr::select(-c("A.y", "B.y", "C.y", "D.y", "E.y", "F.y", "G.y", "H.y", "L.y", "M.y", "N.y"))

length(unique(GO_m_v_c$gene_id))

write.csv(GO_m_v_c, file = "../output/functional_enrichment/Mid_v_Control_GO.csv")
```

Merge high v mid DEGs with GO information
```{r}
## High vs mid treatment comparison 
h_v_m <- read.csv("../output/DEG/High_v_Mid_DEGs.csv") %>%
  rename("gene_id" = "X")

# Merge DEGs and GO information and remove extra columns
GO_h_v_m <- h_v_m %>%
  inner_join(GO_deg, by = "gene_id") %>%
  dplyr::select(-c("A.y", "B.y", "C.y", "D.y", "E.y", "F.y", "G.y", "H.y", "L.y", "M.y", "N.y"))

length(unique(GO_h_v_m$gene_id))

write.csv(GO_h_v_m, file = "../output/functional_enrichment/High_v_Mid_GO.csv")
```

Still need to look at these results more closely. 

I can also run GOSeq on the eggNOG, KEGG and PFAM terms. There are several columns for KEGG information in the `annot` file (KEGG_ko, KEGG_Pathway, KEGG_Module, KEGG_Reaction, KEGG_rclass, KEGG_TC) and I'm not sure what they all mean or which one to use in the analysis. There is also a BRITE column that looks like there is KEGG information in it. 

Select PFAM columns and split so that there is only one PFAM term per row
```{r}
annot_pfam <- annot %>%
  dplyr::select(X.query, PFAMs) %>%
  mutate_all(~if_else(. == "-", NA_character_, .)) %>%
  separate_rows(PFAMs, sep = ",") %>%
  rename(gene_id = X.query,
         PFAM.ID = PFAMs)

# Remove leading and trailing whitespaces
annot_pfam$PFAM.ID <- trimws(annot_pfam$PFAM.ID)

# Replace NAs with unknown
annot_pfam$PFAM.ID <- replace_na(annot_pfam$PFAM.ID, "unknown")

# Set gene and go ids as factors 
annot_pfam$gene_id <- as.factor(annot_pfam$gene_id)
annot_pfam$PFAM.ID <- as.factor(annot_pfam$PFAM.ID)

# Select only unique rows
annot_pfam <- unique(annot_pfam)

# Look at unique gene and go ids 
length(unique(annot_pfam$gene_id))
length(unique(annot_pfam$PFAM.ID))

# Remove rows with unknowns
# annot_GO <- annot_GO %>%
#   filter(GO.ID != "unknown")
```

Remove the annotation information for genes not present in our data 
```{r}
filt_annot_pfam <- subset(annot_pfam, gene_id %in% filt_counts$gene_id)

str(filt_annot_pfam)

# Convert filt_annot_GO from tibble to df 
filt_annot_pfam <-as.data.frame(filt_annot_pfam)
```

Perform GOSeq
```{r}
GO.wall_pfam<-goseq::goseq(DEG.pwf, ID_vector, gene2cat=filt_annot_pfam, method="Wallenius", use_genes_without_cat=T)

pfam <- GO.wall_pfam[order(GO.wall_pfam$over_represented_pvalue),]
write.csv(pfam, file = "../output/functional_enrichment/GOSeq_PFAM_all.csv")

class(GO.wall_pfam)
head(GO.wall_pfam)
tail(GO.wall_pfam)
nrow(GO.wall_pfam)
```

Filter by significantly enriched PFAM terms (p < 0.05)
```{r}
pfam_05 <- pfam %>%
  dplyr::filter(over_represented_pvalue<0.05) %>%
        dplyr::arrange(., over_represented_pvalue)

write.csv(pfam_05, file = "../output/functional_enrichment/GOSeq_PFAM_05.csv")
```

Plot and order by p-value
```{r}
pfam_05_plot <- pfam_05 %>% 
  #drop_na(ontology) %>% 
  #filter(ontology == "MF") %>%
  mutate(category = fct_reorder(category, numDEInCat)) %>%
  #mutate(term = fct_reorder(term)) %>%
  ggplot( aes(x=category, y=numDEInCat) ) +
  geom_segment( aes(x=category ,xend=category, y=0, yend=numDEInCat), color="grey") +
  geom_text(aes(label = over_represented_pvalue), hjust = -1, vjust = 0, size = 2) +
  geom_point(size=1) +
  coord_flip() +
  ylim(0,55) +
  theme(
    panel.grid.minor.y = element_blank(),
    panel.grid.major.y = element_blank(),
    legend.position="bottom"
  ) +
  xlab("") +
  ylab("") +
  theme_bw() + #Set background color 
  theme(panel.border = element_blank(), # Set border
        panel.grid.major = element_blank(), #Set major gridlines
        panel.grid.minor = element_blank(), #Set minor gridlines
        axis.line = element_line(colour = "black"), #Set axes color
        plot.background=element_blank()); pfam_05_plot #Set the plot background #set title attributes

ggsave("../output/functional_enrichment/GOSeq_PFAM_05.pdf", pfam_05_plot, width = 30, height = 10)
ggsave("../output/functional_enrichment/GOSeq_PFAM_05.png", pfam_05_plot, width = 30, height = 10)
```

Still need to look at DEGs in the context of PFAM results and run goseq for KEGG and eggNOG annotation terms


