---
title: "GO MWU input"
author: "Jill Ashey"
date: "2024-07-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r install packages}
#if ("rtracklayer" %in% rownames(installed.packages()) == 'FALSE') BiocManager::install("rtracklayer")
#BiocManager::install("rrvgo")

library(dplyr)
library(ggplot2)
library(rtracklayer)
library(rrvgo)
library(GO.db)
library(tidyr)
library(forcats)
library(scales)
sessionInfo() #list of packages after library-ing these packages
```

### Isolating comparisons and using fold change 

I'm going to isolate comparisons and then use fold change as input to GO MWU rather than a binary of 1 (DE) or 0 (not DE). 

Read in annotation file that will be used for all comparisons. 


#### High v Control 

Read in High v Control DESeq2 results for expressed genes 
```{r}
hvc <- read.csv("../output/DEG/High_v_Control_All.csv") %>%
  dplyr::rename(gene_id = X) %>%
  na.omit() 
```

Match up genes in filtered gene count list to annotation file
```{r}
probes <- hvc$gene_id

# Match genes in filtered count data to genes in annotation file 
probes2annot <- match(probes, annot$gene_id)

# The following is the number of probes without annotation
sum(is.na(probes2annot))
row_nas<-which(is.na(probes2annot))

# View expressed genes that do not have a match in the annotation file 
missing<-as.data.frame(probes[row_nas])
```

There are 7571 genes that were expressed in our data that are not present in the annotation file. 

Reduce annotation file to only contain genes detected in our dataset
```{r}
filtered_annot <- annot[annot$gene_id %in% probes, ]
dim(filtered_annot)
```

The annotation file now only contains genes that were detected in our dataset that have annotation information. 

Format GO terms. For GO_MWU, one line per gene is needed with the GO terms separated by semicolons. 
```{r}
filtered_annot_go <- filtered_annot %>%
  dplyr::select(gene_id, GOs) %>%
  mutate_all(~if_else(. == "-", NA_character_, .)) %>%
  na.omit()

# Change comma to semicolon 
filtered_annot_go$GOs <- gsub(",", ";", filtered_annot_go$GOs)
```

8297 genes were retained in the filtered GO dataset. Write table for GO MWU analysis 
```{r}
write.table(filtered_annot_go, "GO_MWU/High_v_Control_gene_to_go.tab",row.names = FALSE, sep = "\t", quote = FALSE)
```

Filter the counts by the remaining annotations 
```{r}
# Counts 
filt_counts_annot <- filtered_annot_go %>%
  left_join(hvc, by = "gene_id") %>%
  dplyr::select(c("gene_id", "log2FoldChange"))
```

Make GO_MWU input csv file with fold changes 
```{r}
write.csv(filt_counts_annot, "GO_MWU/DE_High_v_Control_LFC.csv", row.names = FALSE, quote = FALSE)
```

#### High v Mid 

Read in High v Mid DESeq2 results for expressed genes 
```{r}
hvm <- read.csv("../output/DEG/High_v_Mid_All.csv") %>%
  dplyr::rename(gene_id = X) %>%
  na.omit() 
```

Match up genes in filtered gene count list to annotation file
```{r}
probes <- hvm$gene_id

# Match genes in filtered count data to genes in annotation file 
probes2annot <- match(probes, annot$gene_id)

# The following is the number of probes without annotation
sum(is.na(probes2annot))
row_nas<-which(is.na(probes2annot))

# View expressed genes that do not have a match in the annotation file 
missing<-as.data.frame(probes[row_nas])
```

There are 9453 genes that were expressed in our data that are not present in the annotation file. 

Reduce annotation file to only contain genes detected in our dataset
```{r}
filtered_annot <- annot[annot$gene_id %in% probes, ]
dim(filtered_annot)
```

The annotation file now only contains genes that were detected in our dataset that have annotation information. 

Format GO terms. For GO_MWU, one line per gene is needed with the GO terms separated by semicolons. 
```{r}
filtered_annot_go <- filtered_annot %>%
  dplyr::select(gene_id, GOs) %>%
  mutate_all(~if_else(. == "-", NA_character_, .)) %>%
  na.omit()

# Change comma to semicolon 
filtered_annot_go$GOs <- gsub(",", ";", filtered_annot_go$GOs)
```

8573 genes were retained in the filtered GO dataset. Write table for GO MWU analysis 
```{r}
write.table(filtered_annot_go, "GO_MWU/High_v_Mid_gene_to_go.tab",row.names = FALSE, sep = "\t", quote = FALSE)
```

Filter the counts by the remaining annotations 
```{r}
# Counts 
filt_counts_annot <- filtered_annot_go %>%
  left_join(hvm, by = "gene_id") %>%
  dplyr::select(c("gene_id", "log2FoldChange"))
```

Make GO_MWU input csv file with fold changes 
```{r}
write.csv(filt_counts_annot, "GO_MWU/DE_High_v_Mid_LFC.csv", row.names = FALSE, quote = FALSE)
```

#### Mid v Control 

Read in Mid v Control DESeq2 results for expressed genes 
```{r}
mvc <- read.csv("../output/DEG/Mid_v_Control_All.csv") %>%
  dplyr::rename(gene_id = X) %>%
  na.omit() 
```

Match up genes in filtered gene count list to annotation file
```{r}
probes <- mvc$gene_id

# Match genes in filtered count data to genes in annotation file 
probes2annot <- match(probes, annot$gene_id)

# The following is the number of probes without annotation
sum(is.na(probes2annot))
row_nas<-which(is.na(probes2annot))

# View expressed genes that do not have a match in the annotation file 
missing<-as.data.frame(probes[row_nas])
```

There are 9765 genes that were expressed in our data that are not present in the annotation file. 

Reduce annotation file to only contain genes detected in our dataset
```{r}
filtered_annot <- annot[annot$gene_id %in% probes, ]
dim(filtered_annot)
```

The annotation file now only contains genes that were detected in our dataset that have annotation information. 

Format GO terms. For GO_MWU, one line per gene is needed with the GO terms separated by semicolons. 
```{r}
filtered_annot_go <- filtered_annot %>%
  dplyr::select(gene_id, GOs) %>%
  mutate_all(~if_else(. == "-", NA_character_, .)) %>%
  na.omit()

# Change comma to semicolon 
filtered_annot_go$GOs <- gsub(",", ";", filtered_annot_go$GOs)
```

8622 genes were retained in the filtered GO dataset. Write table for GO MWU analysis 
```{r}
write.table(filtered_annot_go, "GO_MWU/Mid_v_Control_gene_to_go.tab",row.names = FALSE, sep = "\t", quote = FALSE)
```

Filter the counts by the remaining annotations 
```{r}
# Counts 
filt_counts_annot <- filtered_annot_go %>%
  left_join(mvc, by = "gene_id") %>%
  dplyr::select(c("gene_id", "log2FoldChange"))
```

Make GO_MWU input csv file with fold changes 
```{r}
write.csv(filt_counts_annot, "GO_MWU/DE_Mid_v_Control_LFC.csv", row.names = FALSE, quote = FALSE)
```

