---
title: "Shade levels"
author: "Jill Ashey"
date: "2024-06-06"
output: html_document
---

This script calculates the amount of light in tanks under a shade cloth relative to tanks without a shade cloth. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readxl)
```

Load data
```{r}
shade <- read_excel("../Raw_data/HOBOs/excel/control Shade 2022-06-19 10_01_12 -1000.xlsx")

out <- read_excel("../Raw_data/HOBOs/excel/control Out 2022-06-19 09_58_54 -1000.xlsx")
```

Clean up data 
```{r}
# Remove first two rows and columns 4-7
shade <- shade[-c(1:2), -c(4:7)]
out <- out[-c(1:2), -c(4:7)]

# Rename columns 
colnames(shade) <- c("Date.Time", "Temperature_C", "Light_lux")
colnames(out) <- c("Date.Time", "Temperature_C", "Light_lux")

# Remove NAs
shade <- na.omit(shade)
out <- na.omit(out)
```

Change to numeric
```{r}
shade$Light_lux <- as.numeric(shade$Light_lux)
out$Light_lux <- as.numeric(out$Light_lux)
```


Take mean of light 
```{r}
mean(shade$Light_lux)
mean(out$Light_lux)
```

Mean light of shaded tanks is 2206.019 lux. Mean light of unshaded tanks is 7281.615. 

How much light reduction is the shade cloth doing?
```{r}
(2206.019/7281.615)*100
```

30.29574% of light is coming through to the tanks when the shade cloth is on. 2206.019 is 32.44 umol/m2/s, and 7281.615 is 107.08 umol/m2/s. 




