---
title: "Temperature & pH"
author: "Jill Ashey"
date: "2024-06-06"
output: html_document
---

This script plots temperature and pH before and during the experiment. The experiment was from 7/10/22 - 7/16/22. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

## Before experiment 

In order to know the diurnal temperature and pH conditions, we measured incoming ambient water with hobo loggers for temperature and the apex for pH. We used this data to determine what the mid and high treatments would be in the experiment. Here, I will plot data frome 6/19-6/28. 

Read in data prior to experiment
```{r}
# Define the path to the directory containing the files
path <- "../Raw_data/HOBOs/txt"

# List all files in the directory that contain "2022-06-28" in their names
files <- list.files(path, pattern = "2022-06-28", full.names = TRUE)

# Function to read each file, process data, and add the "tank" column
read_and_process_file <- function(file) {
  # Read the file into a data frame
  df <- read.delim(file, header = F)  # Use read_csv for CSV files; adjust accordingly for other formats
  
  # Remove first two rows and columns 4-8
  df <- df[-c(1:2), -c(4:8)]
  
  # Rename columns
  colnames(df) <- c("Date.Time", "Temperature_C", "Light_lux")
  
  # Remove NAs
  df <- df %>%
    mutate(across(everything(), ~na_if(., ""))) %>%
    na.omit()
  
  # Replace commas with periods in temp and light columns 
  df$Temperature_C <- gsub(",", ".", df$Temperature_C)
  df$Light_lux <- gsub(",", ".", df$Light_lux)
  
  # Extract the first 5 characters from the file name for the "tank" column
  tank_name <- substr(basename(file), 1, 5)
  
  # Add the "tank" column to the data frame
  df <- df %>%
    mutate(tank = tank_name)
  
  return(df)
}

# Use purrr::map_df to read all files and bind them into one data frame
combined_df <- map_df(files, read_and_process_file)
```

Make temperature and light variables numeric 
```{r}
combined_df$Temperature_C <- as.numeric(combined_df$Temperature_C)
combined_df$Light_lux <- as.numeric(combined_df$Light_lux)
```

Convert Date.Time to POSIXct and add Time column
```{r}
combined_df <- combined_df %>%
  mutate(Date.Time = as.POSIXct(Date.Time, format = "%Y-%m-%d %H:%M:%S")) %>%
  mutate(Time = format(Date.Time, "%H:%M"))
```

```{r}
head(combined_df)
```





Group by Time and calculate means 
```{r}
# Group by Time and calculate the mean of Temperature_C and Light_lux
time_averages <- combined_df %>%
  group_by(Time) %>%
  summarise(
    Avg_Temperature_C = mean(Temperature_C, na.rm = F),
    sd_Temperature_C = sd(Temperature_C, na.rm = TRUE) / sqrt(n()),
    Avg_Light_lux = mean(Light_lux, na.rm = F)/60,
    sd_Light_lux = sd(Light_lux, na.rm = TRUE) / sqrt(n())) %>%
  mutate(upper_temp = Avg_Temperature_C+sd_Temperature_C, lower_temp = Avg_Temperature_C-sd_Temperature_C) %>%
  na.omit() %>%
  mutate(Time = hms::parse_hm(Time))
```

Plot 
```{r}
ggplot(time_averages, aes(x = Time, y = Avg_Temperature_C)) +
    #geom_ribbon(aes(ymin = lower_temp, ymax = upper_temp), alpha = 0.4, show.legend=FALSE) +
  geom_smooth(method="loess", span=0.5, se=T, alpha=0.5, show.legend=TRUE) +
  #geom_line(color = "blue", linewidth = 0.5) +  # Adjust the size parameter for thinner lines
  #geom_ribbon(aes(ymin = lower_temp, ymax = upper_temp), alpha = 0.05, fill = "blue") +
  labs(x = "Time",
       y = "Average Temperature (째C)") +
  theme_minimal()
```

Plot part 2 - can't figure out how to plot the upper and lower 
```{r}
ggplot(time_averages, aes(x = Time, y = Avg_Temperature_C)) +
  geom_ribbon(aes(ymin = lower_temp, ymax = upper_temp), alpha = 1, show.legend=FALSE) +
  geom_smooth(method="loess", span=0.5, se=F, alpha=0.5, show.legend=TRUE) +
  #geom_ribbon(aes(ymin = lower_temp, ymax = upper_temp), alpha = 0.05, fill = "blue") +
  labs(x = "Time",
       y = "Average Temperature (째C)") +
  theme_minimal()
```






Read in pH data from apex 
```{r}
ph_data <- read.delim("../Raw_data/Apex/Data_log_control_beforeExp.xlsx.csv", sep = ";")
```

Convert Date.Time to POSIXct and add Time column
```{r}
ph_data <- ph_data %>%
  mutate(record.date = as.POSIXct(record.date, format = "%Y-%m-%d %H:%M:%S")) %>%
  mutate(Time = format(record.date, "%H:%M"))
```

Filter so that there is only data from 6/19-6/28
```{r}
ph_data_filt <- ph_data %>%
  dplyr::filter(record.date < as.POSIXct("2022-06-28 23:59:00"))
```

Group by Time and calculate means 
```{r}
time_averages <- ph_data_filt %>%
  group_by(Time) %>%
  summarise(
    Avg_ph = mean(record.probe.value, na.rm = F),
    sd_ph = sd(record.probe.value, na.rm = TRUE) / sqrt(n()))%>%
  mutate(upper_ph = Avg_ph+sd_ph, lower_ph = Avg_ph-sd_ph) %>%
  na.omit() %>%
  mutate(Time = hms::parse_hm(Time))
```

Plot 
```{r}
ggplot(time_averages, aes(x = Time, y = Avg_ph)) +
    #geom_ribbon(aes(ymin = lower_temp, ymax = upper_temp), alpha = 0.4, show.legend=FALSE) +
  geom_smooth(method="loess", span=0.5, se=T, alpha=0.5, show.legend=TRUE) +
  #geom_line(color = "blue", linewidth = 0.5) +  # Adjust the size parameter for thinner lines
  #geom_ribbon(aes(ymin = lower_temp, ymax = upper_temp), alpha = 0.05, fill = "blue") +
  labs(x = "Time",
       y = "Average pH (NBS)") +
  theme_minimal()
```

Join temperature and pH info
```{r}
before_exp <- time_averages %>%
  inner_join(time_averages_ph, by = "Time")

# Transform ph
before_exp <- before_exp %>%
  mutate(ph_transform = Avg_ph*(1+0.1*(Avg_Temperature_C-25)))
```

Plot together???
```{r}
ggplot(before_exp, aes(x = Time)) +
  geom_line(aes(y = Avg_Temperature_C, color = "Avg_Temperature_C")) +
  geom_line(aes(y = ph_transform, color = "Avg_ph")) +
  scale_y_continuous(
    name = "Average Temperature (째C)",  # Left y-axis label
    sec.axis = sec_axis(~., name = "Average pH")  # Right y-axis label
  ) +
  labs(
    x = "Time",
    color = NULL
  ) +
  scale_color_manual(
    name = "Variables",
    values = c("Avg_Temperature_C" = "blue", "Avg_ph" = "red")
  ) +
  theme_minimal()
```


## During experiment 

The following data is from the experiment from 7/10/22-7/16/22. 

Read in data prior to experiment
```{r}
# Define the path to the directory containing the files
path <- "../Raw_data/HOBOs/txt"

# List all files in the directory that contain "2022-06-28" in their names
files <- list.files(path, pattern = "2022-07-16", full.names = TRUE)

# Function to read each file, process data, and add the "tank" column
read_and_process_file <- function(file) {
  # Read the file into a data frame
  df <- read.delim(file, header = F)  # Use read_csv for CSV files; adjust accordingly for other formats
  
  # Remove first two rows and columns 4-8
  df <- df[-c(1:2), -c(4:8)]
  
  # Rename columns
  colnames(df) <- c("Date.Time", "Temperature_C", "Light_lux")
  
  # Remove NAs
  df <- df %>%
    mutate(across(everything(), ~na_if(., ""))) %>%
    na.omit()
  
  # Replace commas with periods in temp and light columns 
  df$Temperature_C <- gsub(",", ".", df$Temperature_C)
  df$Light_lux <- gsub(",", ".", df$Light_lux)
  
  # Extract the first 5 characters from the file name for the "tank" column
  tank_name <- substr(basename(file), 1, 5)
  
  # Add the "tank" column to the data frame
  df <- df %>%
    mutate(tank = tank_name)
  
  return(df)
}

# Use purrr::map_df to read all files and bind them into one data frame
combined_df <- map_df(files, read_and_process_file)
```

Make temperature and light variables numeric 
```{r}
combined_df$Temperature_C <- as.numeric(combined_df$Temperature_C)
combined_df$Light_lux <- as.numeric(combined_df$Light_lux)
```

Convert Date.Time to POSIXct 
```{r}
combined_df <- combined_df %>%
  mutate(Date.Time = as.POSIXct(Date.Time, format = "%Y-%m-%d %H:%M:%S")) 
```

Filter so that Date.Time starts at 7/10
```{r}
during_exp <- combined_df %>%
  dplyr::filter(Date.Time > as.POSIXct("2022-07-10 08:00:00"))
```

Make Treatment column and remove NAs
```{r}
during_exp <- during_exp %>%
  mutate(Treatment = case_when(
    grepl("C", tank) ~ "Control",
    grepl("M", tank) ~ "Mid",
    grepl("H", tank) ~ "High",
    TRUE ~ NA_character_  # If none of the conditions match, return NA
  )) %>%
  na.omit()
```

```{r}
head(during_exp)
```








Group by Time and calculate means 
```{r}
# Group by Time and calculate the mean of Temperature_C and Light_lux
during_exp_avg <- during_exp %>%
  group_by(Treatment, Date.Time) %>%
  summarise(
    Avg_Temperature_C = mean(Temperature_C, na.rm = F),
    sd_Temperature_C = sd(Temperature_C, na.rm = TRUE) / sqrt(n()),
    Avg_Light_lux = mean(Light_lux, na.rm = F)/60,
    sd_Light_lux = sd(Light_lux, na.rm = TRUE) / sqrt(n())) %>%
  mutate(upper_temp = Avg_Temperature_C+sd_Temperature_C, lower_temp = Avg_Temperature_C-sd_Temperature_C) %>%
  na.omit() 
```










```{r}
blah <- ggplot(during_exp, aes(x = Date.Time, y = Temperature_C, color = Treatment)) +
  geom_smooth(method="loess", span=0.5, se=F, alpha=0.5, show.legend=TRUE) +
  #geom_line()
  labs(
    x = "Date",
    y = "Temperature (째C)"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  ); blah
  
#ggsave("../output/blah.pdf", blah, width = 20, height = 11)

```


Need to troubleshoot and save plots
